\singlespacing{\textbf{Efficient pollution abatement in electricity markets with intermittent renewable energy}}

### 2019-04-28 documents/draft.tex
Update draft
- add further discussion of results
- update data section, mentioning deflation
- add figures for discussion

\section{Data}

We collect monthly data from the \cite{EIANetgen} on retail electricity prices and consumption for each state in the US from 2011 to 2018. Additionally, also from the EIA, we obtain data on the average cost of coal for electricity generation for each state and month.\footnote{The coal price data set contains a large number of missing values due to privacy concerns; however, we do not expect that these missing values are correlated with the data itself.} 
Finally, we collect data on HDDs and CDDs from the NOAA's \citet{CPC} for the same panel. Then, we merge these three data sets and trim 1\% of outliers for a total of $826$ observations for each month and state. We use this preliminary data set to construct the data required for our regressions. That is, each observation in our estimation equation belongs to a set $(t,s,i)$ consisting of two time periods and a state. Hence, we construct each row in our regression data set using unique combinations of $t,s$ where $t \neq s$ for each state $i$. Due to the number of potential combinations and thus observations, we restrict our data set to a random sample of 9000 observations.\footnote{We reran our regressions with several different samples of 9000 and found that our estimated coefficients did not significantly change; hence we believe this sample size is sufficient.} 
All in all, each observation in our regression data set consists the following variables: state ($i$), date 1 ($t$), date 2 ($s$), the log difference in electricity consumption between month 1 and month 2, the log difference in the price of electricity, the log difference in the price of coal, the number of CDDs for each date, the number of HDDs for each date, and the difference (in months) between dates 1 and 2.

@@

\begin{table}[!htbp] \centering 
	\caption{OLS Regression Results}
	\label{table:1} 
	\small
	\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
		\\[-4ex]\hline  
		\hline \\[-1.8ex] 
		& \multicolumn{6}{c}{\textit{Dependent variable:} $\ln (Z_{ t, i} / Z_{ s, i})$} \\ [0.5ex]
		\cline{2-7} 
		\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ [0.5ex]
		\hline \\[-1.8ex] 
		$-\ln (P_{t,i} / P_{s,i})$ & 0.751$^{***}$ & 0.481$^{***}$ & 0.607$^{***}$ & 0.750$^{***}$ & 0.481$^{**}$ & 0.607$^{***}$ \\ 
		& (0.030) & (0.026) & (0.026) & (0.224) & (0.172) & (0.169) \\ 
		& & & & & & \\ 
		$\Delta_{t,s}$ &  &  & 0.001$^{***}$ &  &  & 0.001$^{***}$ \\ 
		&  &  & (0.0001) &  &  & (0.0002) \\ 
		& & & & & & \\ 
		CDD$_t$ &  & 0.990$^{***}$ & 1.003$^{***}$ &  & 0.990$^{***}$ & 1.002$^{***}$ \\ 
		$\quad(\times 1000^{-1})$&  & (0.016) & (0.016) &  & (0.087) & (0.086) \\ 
		& & & & & & \\ 
		CDD$_s$ &  & $-$0.999$^{***}$ & $-$1.011$^{***}$ &  & $-$0.999$^{***}$ & $-$1.012$^{***}$ \\ 
		$\quad(\times 1000^{-1})$&  & (0.016) & (0.016) &  & (0.087) & (0.085) \\ 
		& & & & & & \\ 
		HDD$_t$ &  & 0.306$^{***}$ & 0.294$^{***}$ &  & 0.307$^{***}$ & 0.295$^{***}$ \\ 
		$\quad(\times 1000^{-1})$&  & (0.005) & (0.005) &  & (0.030) & (0.030) \\ 
		& & & & & & \\ 
		HDD$_s$ &  & $-$0.309$^{***}$ & $-$0.296$^{***}$ &  & $-$0.308$^{***}$ & $-$0.295$^{***}$ \\ 
		$\quad(\times 1000^{-1})$&  & (0.005) & (0.005) &  & (0.029) & (0.029) \\ 
		& & & & & & \\ 
		Intercept & 0.001 & 0.002 & 0.002 &  &  &  \\ 
		& (0.003) & (0.006) & (0.006) &  &  &  \\ 
		& & & & & & \\  [0.9ex]
		\hline \\[-1.8ex] 
		State FEs &   &   &   & Yes & Yes & Yes \\ 
		Observations & 9,000 & 9,000 & 9,000 & 9,000 & 9,000 & 9,000 \\ 
		R$^{2}$ & 0.070 & 0.580 & 0.596 & 0.070 & 0.580 & 0.596 \\ 
		Adjusted R$^{2}$ & 0.070 & 0.579 & 0.596 & 0.065 & 0.577 & 0.594 \\ [0.5ex]
		\hline 
		\hline \\[-1.8ex] 
		\multicolumn{7}{@{}p{40em}@{}}{\textit{Note: } The sample covers all 50 US states from 2011 to 2018; outliers are removed by trimming 1\% of each variable except $\Delta_{t,s}$. The unit of observation is a set $(t,s,i)$ where $t \neq s$ are months and $i$ is a state; as discussed in the Data section, we take a random sample of 9000 observations from the data. The coefficient on $\ln (P_{t,i} / P_{s,i})$ is an estimate of $-\sigma$. The variable $\Delta_{t,s}$ is the difference in months between periods $t$ and $s$. CDD$_t$ and HDD$_t$ refer to the total number of heating and cooling degree days in month $t$. Robust standard errors are reported in parentheses. *p$\textless$0.05, **p$\textless$0.01, ***p$\textless$0.001}  \\ 
	\end{tabular} 
\end{table}

\begin{table}[!htbp] \centering 
	\caption{IV (2SLS) Regression Results}
	\label{table:2} 
	\small
	\begin{tabular}{@{\extracolsep{4pt}}lcccccc} 
		\\[-4ex]\hline  
		\hline \\[-1.6ex] 
		& \multicolumn{3}{c}{First-Stage} & \multicolumn{3}{c}{Second-Stage} \\ [0.5ex]
		& \multicolumn{3}{c}{\textit{Dep. Variable:} $\ln (P_{t,i} / P_{s,i})$ } & \multicolumn{3}{c}{\textit{Dep. Variable:}  $\ln (Z_{ t, i} / Z_{ s, i})$}\\ [0.5ex]
		\cmidrule(lr){2-4} \cmidrule(lr){5-7}\\[-2.2ex] 
		& (A.1) & (B.1) & (C.1) & (A.2) & (B.2) & (C.2)\\ [0.5ex]
		\hline \\[-1.8ex] 
		$ \ln (C_{t,i} / C_{s,i})$ & $-$0.060$^{***}$ & 0.004$^{*}$ & 0.004$^{*}$ &  &  &  \\ 
		& (0.002) & (0.002) & (0.002) &  &  &  \\ 
		& & & & & & \\ 
		$-\ln (P_{t,i} / P_{s,i})$ &  &  &  & 0.711$^{***}$ & 13.987 & 11.430$^{*}$ \\ 
		&  &  &  & (0.106) & (8.384) & (5.330) \\ 
		& & & & & & \\ 
		$\Delta_{t,s}$  &  & 0.001$^{***}$ & 0.001$^{***}$ &  &  & 0.007$^{*}$ \\ 
		&  & (0.00002) & (0.00002) &  &  & (0.003) \\ 
		& & & & & & \\ 
		CDD$_t$  &  & 0.151$^{***}$ & 0.151$^{***}$ &  & 3.064$^{*}$ & 2.620$^{**}$ \\ 
		$\quad(\times 1000^{-1})$&  & (0.007) & (0.007) &  & (1.288) & (0.798) \\ 
		& & & & & & \\ 
		CDD$_s$  &  & $-$0.149$^{***}$ & $-$0.149$^{***}$ &  & $-$3.034$^{*}$ & $-$2.609$^{***}$ \\ 
		$\quad(\times 1000^{-1})$&  & (0.007) & (0.007) &  & (1.264) & (0.788) \\ 
		& & & & & & \\ 
		HDD$_t$  &  & $-$0.057$^{***}$ & $-$0.057$^{***}$ &  & $-$0.407 & $-$0.311 \\ 
		$\quad(\times 1000^{-1})$&  & (0.003) & (0.003) &  & (0.447) & (0.301) \\ 
		& & & & & & \\ 
		HDD$_s$  &  & 0.058$^{***}$ & 0.058$^{***}$ &  & 0.423 & 0.324 \\ 
		$\quad(\times 1000^{-1})$&  & (0.003) & (0.003) &  & (0.458) & (0.308) \\ [0.9ex]
		\hline \\[-1.8ex] 
		State FEs &  Yes & Yes  & Yes  & Yes & Yes & Yes \\ 
		Observations & 9,000 & 9,000 & 9,000 & 9,000 & 9,000 & 9,000 \\ 
		R$^{2}$          & 0.066 & 0.441 & 0.441 & & &  \\ 
		Adjusted R$^{2}$ & 0.066 & 0.441 & 0.441 & & &  \\ 
		F Statistic & 678$^{***}$ & 1,213$^{***}$ & 1,197$^{***}$  &  &  &  \\ 
		\hline 
		\hline \\[-1.8ex] 
		\multicolumn{7}{@{}p{39.2em}@{}}{\textit{Note: } The log difference in coal price between period $t$ and $s$, $ \ln (C_{t,i} / C_{s,i})$, is used as an instrument in these regressions. The sample covers all 50 US states from 2011 to 2018; outliers are removed by trimming 1\% of each variable except $\Delta_{t,s}$. The unit of observation is a set $(t,s,i)$ where $t \neq s$ are months and $i$ is a state; as discussed in the Data section, we take a random sample of 9000 observations from the data. The coefficient on $\ln (P_{t,i} / P_{s,i})$ is an estimate of $-\sigma$. The variable $\Delta_{t,s}$ is the difference in months between periods $t$ and $s$. CDD$_t$ and HDD$_t$ refer to the total number of heating and cooling degree days in month $t$. Robust standard errors are reported in parentheses. *p$\textless$0.05, **p$\textless$0.01, ***p$\textless$0.001}  \\ 
	\end{tabular} 
\end{table}

\begin{table}[!htbp] \centering 
	\caption{Partially Linear IV Regression Results}
	\label{table:3} 
	\small
	\begin{tabular}{@{\extracolsep{4em}}lccc} 
		\\[-4ex]\hline  
		\hline \\[-1.8ex] 
		& \multicolumn{3}{c}{\textit{Instrument: $ \ln (C_{t,i} / C_{s,i})$}} \\ 
		\cline{2-4} 
		%\\[-1.8ex] & ln\_load\_rel \textasciitilde ln\_price\_rel & ln\_load\_rel \textasciitilde (CDD\_1) & ln\_load\_rel \textasciitilde time\_diff + (CDD\_1) \\ 
		\\[-1.8ex] & (1) & (2) & (3)\\ [0.5ex]
		\hline \\[-1.8ex] 
		$\hat{\sigma} $ & 0.5676$^{***}$ & 0.4511$^{***}$ & 0.4129$^{***}$ \\ 
		& (0.1036) & (0.0042) & (0.0085) \\ [0.9ex]
		\hline \\[-1.8ex] 
		Time Control &   &   & Yes  \\ 
		Degree Day Controls &   & Yes  & Yes  \\ 
		Observations & 9,000 & 9,000 & 9,000 \\
		\hline 
		\hline \\[-1.8ex] 
		\multicolumn{4}{@{}p{36em}@{}}{\textit{Note: } The log difference in coal price between period $t$ and $s$, $ \ln (C_{t,i} / C_{s,i})$, is used as an instrument in these regressions. The sample covers all 50 US states from 2011 to 2018; outliers are removed by trimming 1\% of each variable except $\Delta_{t,s}$. The unit of observation is a set $(t,s,i)$ where $t \neq s$ are months and $i$ is a state; as discussed in the Data section, we take a random sample of 9000 observations from the data. The estimation procedure is described in the appendix.  Robust standard errors are reported in parentheses. *p$\textless$0.05, **p$\textless$0.01, ***p$\textless$0.001}  \\ 
	\end{tabular} 
\end{table}

@@

\subsubsection{The Relationship between Renewable and Fossil Energy}

@@

Firstly, we let technology 1 be coal power and technology 2 be solar power; again, period $t$ represents peak hours and period $s$ represents off-peak hours. We assume that, holding prices equal, consumers prefer that 60\% of their energy arrive in period $t$ and the remaining 40\% arrive in period $s$; that is, we have $\alpha_t = 0.6$ and $\alpha_s = 0.4$. Additionally, we normalize the units of the quantities to a MWh basis. Hence, $\xi$ represents the percent of capacity utilized in each period; we assume coal uses 100\% of its capacity in both periods, while solar can access 100\% during peak hours and only 10\% during the off-peak. Finally, we set cost parameters (given in  \$ per MWh)  equal to LCOE estimates for 2023 from the \citet{EIALCOE}; specifically, we use estimates for "Solar PV" and "Coal with 30\% CCS." In total, we have the following parameters.

@@

\begin{table}[t]
	\caption*{\textbf{Figure 1:} Elasticity of Substition Between Solar and Coal} 
	\label{fig:eosnum}
	\footnotesize
	\begin{tabular}{@{\extracolsep{0em}}c}
		\includegraphics[width=1\linewidth]{../code/matlab/simulation_april/example_1_fig_1} \\
		\multicolumn{1}{@{\hspace{0.2in}}p{5.9in}@{}}{ \textit{Note: } The y-axis of the first plot is equivalent to $\log(X_1/X_2)$ and the x-axis of both plots is equivalent to  $\log(c_2/c_1)$. The legend in the upper subplot also applies to the lower subplot. These results were obtained using the following parameters: $\alpha_t = 0.6$, $\alpha_s = 0.4$, $\xi_1 = (1, \, 1)$, $\xi_2 = (1, \, 0.1)$, $c_1 = 104.3$, $c_2 = 60$. We plot results for various values of the intertemporal elasticity of substition for electricity consumption $\sigma$, which is used in the consumer's CES utility function. In order to numerically estimate these relationships, we first found the optimal quantities of $X$ over a range of prices $c_1^* \in (0.5\, c_1, 2 \,c_1)$. Then, we obtained estimates of the elasticity of substition by numerically differentiating $\log(X_1/X_2)$ with respect to $-log(c_1, c_2)$. That is, the elasticity of substitution between technology 1 and 2 is given by the slope of the upper subplot, and it is graphed in the lower subplot.   }     \\  
	\end{tabular}
\end{table}

We plot our numerical estimates of $e_{solar, coal}$, the elasticity of substitution between solar and coal power, in \hyperref[fig:eosnum]{Figure 1}. These results were generated by computing the optimal quantities of each technology given a range of prices. Specifically, we held the price of solar constant and varied the price of coal from 50\% to 200\% of its original parametrized price. Then, we numerical differentiated $\log(X_1/X_2)$ with respect to $\log(c_2/c_1)$ to estimate the elasticity of substitution between solar and coal. This numerically computed elasticity of substitution $e$ on the production side is shown in the lower subplot. We repeat this process with different values of $\sigma$, the elasticity of substitution in the consumer's utility function. Recall that $\sigma < 1$ implies that  electricity consumption in different periods are complements, $\sigma = 1$ implies to a Cobb-Douglas relationship, and $\sigma > 1$ implies that electricity consumption in different periods are substitutes. Our results strongly suggest that a CES production structure does not accurately model intermittency.

Firstly, note that the elasticity of substitution becomes increasingly non-linear as $\sigma$ rises. This is somewhat intuitive; when energy consumption can be substituted across periods, the market is more sensitive to the cost efficiency of technologies and less sensitive to when they produce. For instance, suppose we had two technologies that each produced in a different period; if consumers are less willing to substitute consumption across periods, both technologies are likely to be  employed, On the other hand, when consumers are open to substitution (implying large $\sigma$), they will simply consume the majority of their electricity in whichever period it is most cost effective to do so. Consequently, price plays bigger role than intermittency as $\sigma$ rises. This is further evident in the lower subplots; note that the numerical estimates of  $e$ rise as $\sigma$ rises. When consumers are less concerned about when they receive their energy, they are far more open to substituting between coal and solar power.

Second, we find that $e$ approaches a "U" shape as $\sigma$ rises....

...Like \cite{HH} we find an S-shaped adoption curve for renewable energy driven by intermittency.\footnote{\citet{HH} reference \citet{Geroski2000} who provides a survey of the literature on adoption of new technologies. He finds that S-shaped adoption curves are often generated by information cascades. On the other hand, our model and that of Helm and Mier find S-shaped curves as a result of intermittency slowing down adoption.}

>>>>

\subsection{Data}

We collect monthly data from the \cite{EIANetgen} on retail electricity prices and consumption for each state in the US from 2011 to 2018. Additionally, also from the EIA, we obtain data on the average cost of coal for electricity generation for each state and month.\footnote{The coal price data set contains a large number of missing values due to privacy concerns; however, we do not expect that these missing values are correlated with the data itself.} We deflated both electricity and coal prices over time using the PCECP Index provided by the  \cite{USBEA}. 
Finally, we collect data on HDDs and CDDs from the NOAA's \citet{CPC} for the same panel. Then, we merge these three data sets and trim 1\% of outliers for a total of $826$ observations for each month and state. We use this preliminary data set to construct the data required for our regressions. That is, each observation in our estimation equation belongs to a set $(t,s,i)$ consisting of two time periods and a state. Hence, we construct each row in our regression data set using unique combinations of $t,s$ where $t \neq s$ for each state $i$. Due to the number of potential combinations and thus observations, we restrict our data set to a random sample of 9000 observations.\footnote{We reran our regressions with several different samples of 9000 and found that our estimated coefficients did not significantly change; hence we believe this sample size is sufficient.} 
All in all, each observation in our regression data set consists the following variables: state ($i$), date 1 ($t$), date 2 ($s$), the log difference in electricity consumption between month 1 and month 2, the log difference in the price of electricity, the log difference in the price of coal, the number of CDDs for each date, the number of HDDs for each date, and the difference (in months) between dates 1 and 2.

@@

\begin{table}[!t] \centering 
	\caption{OLS Regression Results}
	\label{table:1} 
	\small
	\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
		\\[-4ex]\hline  
		\hline \\[-1.8ex] 
		& \multicolumn{6}{c}{\textit{Dependent variable:} $\ln (Z_{ t, i} / Z_{ s, i})$} \\ [0.5ex]
		\cline{2-7} 
		\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ [0.5ex]
		\hline \\[-1.8ex] 
		$-\ln (P_{t,i} / P_{s,i})$ & 0.751$^{***}$ & 0.481$^{***}$ & 0.607$^{***}$ & 0.750$^{***}$ & 0.481$^{**}$ & 0.607$^{***}$ \\ 
		& (0.030) & (0.026) & (0.026) & (0.224) & (0.172) & (0.169) \\ 
		& & & & & & \\ 
		$\Delta_{t,s}$ &  &  & 0.001$^{***}$ &  &  & 0.001$^{***}$ \\ 
		&  &  & (0.0001) &  &  & (0.0002) \\ 
		& & & & & & \\ 
		CDD$_t$ &  & 0.990$^{***}$ & 1.003$^{***}$ &  & 0.990$^{***}$ & 1.002$^{***}$ \\ 
		$\quad(\times 1000^{-1})$&  & (0.016) & (0.016) &  & (0.087) & (0.086) \\ 
		& & & & & & \\ 
		CDD$_s$ &  & $-$0.999$^{***}$ & $-$1.011$^{***}$ &  & $-$0.999$^{***}$ & $-$1.012$^{***}$ \\ 
		$\quad(\times 1000^{-1})$&  & (0.016) & (0.016) &  & (0.087) & (0.085) \\ 
		& & & & & & \\ 
		HDD$_t$ &  & 0.306$^{***}$ & 0.294$^{***}$ &  & 0.307$^{***}$ & 0.295$^{***}$ \\ 
		$\quad(\times 1000^{-1})$&  & (0.005) & (0.005) &  & (0.030) & (0.030) \\ 
		& & & & & & \\ 
		HDD$_s$ &  & $-$0.309$^{***}$ & $-$0.296$^{***}$ &  & $-$0.308$^{***}$ & $-$0.295$^{***}$ \\ 
		$\quad(\times 1000^{-1})$&  & (0.005) & (0.005) &  & (0.029) & (0.029) \\ 
		& & & & & & \\ 
		Intercept & 0.001 & 0.002 & 0.002 &  &  &  \\ 
		& (0.003) & (0.006) & (0.006) &  &  &  \\ 
		& & & & & & \\  [0.9ex]
		\hline \\[-1.8ex] 
		State FEs &   &   &   & Yes & Yes & Yes \\ 
		Observations & 9,000 & 9,000 & 9,000 & 9,000 & 9,000 & 9,000 \\ 
		R$^{2}$ & 0.070 & 0.580 & 0.596 & 0.070 & 0.580 & 0.596 \\ 
		Adjusted R$^{2}$ & 0.070 & 0.579 & 0.596 & 0.065 & 0.577 & 0.594 \\ [0.5ex]
		\hline 
		\hline \\[-1.8ex] 
		\multicolumn{7}{@{}p{40em}@{}}{\textit{Note: } The sample covers all 50 US states from 2011 to 2018; outliers are removed by trimming 1\% of each variable except $\Delta_{t,s}$. The unit of observation is a set $(t,s,i)$ where $t \neq s$ are months and $i$ is a state; as discussed in the Data section, we take a random sample of 9000 observations from the data. The coefficient on $\ln (P_{t,i} / P_{s,i})$ is an estimate of $-\sigma$. The variable $\Delta_{t,s}$ is the difference in months between periods $t$ and $s$. CDD$_t$ and HDD$_t$ refer to the total number of heating and cooling degree days in month $t$. Robust standard errors are reported in parentheses. *p$\textless$0.05, **p$\textless$0.01, ***p$\textless$0.001}  \\ 
	\end{tabular} 
\end{table}

\begin{table}[!t] \centering 
	\caption{IV (2SLS) Regression Results}
	\label{table:2} 
	\small
	\begin{tabular}{@{\extracolsep{4pt}}lcccccc} 
		\\[-4ex]\hline  
		\hline \\[-1.6ex] 
		& \multicolumn{3}{c}{First-Stage} & \multicolumn{3}{c}{Second-Stage} \\ [0.5ex]
		& \multicolumn{3}{c}{\textit{Dep. Variable:} $\ln (P_{t,i} / P_{s,i})$ } & \multicolumn{3}{c}{\textit{Dep. Variable:}  $\ln (Z_{ t, i} / Z_{ s, i})$}\\ [0.5ex]
		\cmidrule(lr){2-4} \cmidrule(lr){5-7}\\[-2.2ex] 
		& (A.1) & (B.1) & (C.1) & (A.2) & (B.2) & (C.2)\\ [0.5ex]
		\hline \\[-1.8ex] 
		$ \ln (C_{t,i} / C_{s,i})$ & $-$0.060$^{***}$ & 0.004$^{*}$ & 0.004$^{*}$ &  &  &  \\ 
		& (0.002) & (0.002) & (0.002) &  &  &  \\ 
		& & & & & & \\ 
		$-\ln (P_{t,i} / P_{s,i})$ &  &  &  & 0.711$^{***}$ & 13.987 & 11.430$^{*}$ \\ 
		&  &  &  & (0.106) & (8.384) & (5.330) \\ 
		& & & & & & \\ 
		$\Delta_{t,s}$  &  & 0.001$^{***}$ & 0.001$^{***}$ &  &  & 0.007$^{*}$ \\ 
		&  & (0.00002) & (0.00002) &  &  & (0.003) \\ 
		& & & & & & \\ 
		CDD$_t$  &  & 0.151$^{***}$ & 0.151$^{***}$ &  & 3.064$^{*}$ & 2.620$^{**}$ \\ 
		$\quad(\times 1000^{-1})$&  & (0.007) & (0.007) &  & (1.288) & (0.798) \\ 
		& & & & & & \\ 
		CDD$_s$  &  & $-$0.149$^{***}$ & $-$0.149$^{***}$ &  & $-$3.034$^{*}$ & $-$2.609$^{***}$ \\ 
		$\quad(\times 1000^{-1})$&  & (0.007) & (0.007) &  & (1.264) & (0.788) \\ 
		& & & & & & \\ 
		HDD$_t$  &  & $-$0.057$^{***}$ & $-$0.057$^{***}$ &  & $-$0.407 & $-$0.311 \\ 
		$\quad(\times 1000^{-1})$&  & (0.003) & (0.003) &  & (0.447) & (0.301) \\ 
		& & & & & & \\ 
		HDD$_s$  &  & 0.058$^{***}$ & 0.058$^{***}$ &  & 0.423 & 0.324 \\ 
		$\quad(\times 1000^{-1})$&  & (0.003) & (0.003) &  & (0.458) & (0.308) \\ [0.9ex]
		\hline \\[-1.8ex] 
		State FEs &  Yes & Yes  & Yes  & Yes & Yes & Yes \\ 
		Observations & 9,000 & 9,000 & 9,000 & 9,000 & 9,000 & 9,000 \\ 
		R$^{2}$          & 0.066 & 0.441 & 0.441 & & &  \\ 
		Adjusted R$^{2}$ & 0.066 & 0.441 & 0.441 & & &  \\ 
		F Statistic & 678$^{***}$ & 1,213$^{***}$ & 1,197$^{***}$  &  &  &  \\ 
		\hline 
		\hline \\[-1.8ex] 
		\multicolumn{7}{@{}p{39.2em}@{}}{\textit{Note: } The log difference in coal price between period $t$ and $s$, $ \ln (C_{t,i} / C_{s,i})$, is used as an instrument in these regressions. The sample covers all 50 US states from 2011 to 2018; outliers are removed by trimming 1\% of each variable except $\Delta_{t,s}$. The unit of observation is a set $(t,s,i)$ where $t \neq s$ are months and $i$ is a state; as discussed in the Data section, we take a random sample of 9000 observations from the data. The coefficient on $\ln (P_{t,i} / P_{s,i})$ is an estimate of $-\sigma$. The variable $\Delta_{t,s}$ is the difference in months between periods $t$ and $s$. CDD$_t$ and HDD$_t$ refer to the total number of heating and cooling degree days in month $t$. Robust standard errors are reported in parentheses. *p$\textless$0.05, **p$\textless$0.01, ***p$\textless$0.001}  \\ 
	\end{tabular} 
\end{table}

\begin{table}[!t] \centering 
	\caption{Partially Linear IV Regression Results}
	\label{table:3} 
	\small
	\begin{tabular}{@{\extracolsep{4em}}lccc} 
		\\[-4ex]\hline  
		\hline \\[-1.8ex] 
		& \multicolumn{3}{c}{\textit{Instrument: $ \ln (C_{t,i} / C_{s,i})$}} \\ 
		\cline{2-4} 
		%\\[-1.8ex] & ln\_load\_rel \textasciitilde ln\_price\_rel & ln\_load\_rel \textasciitilde (CDD\_1) & ln\_load\_rel \textasciitilde time\_diff + (CDD\_1) \\ 
		\\[-1.8ex] & (1) & (2) & (3)\\ [0.5ex]
		\hline \\[-1.8ex] 
		$\hat{\sigma} $ & 0.5676$^{***}$ & 0.4511$^{***}$ & 0.4129$^{***}$ \\ 
		& (0.1036) & (0.0042) & (0.0085) \\ [0.9ex]
		\hline \\[-1.8ex] 
		Time Control &   &   & Yes  \\ 
		Degree Day Controls &   & Yes  & Yes  \\ 
		Observations & 9,000 & 9,000 & 9,000 \\
		\hline 
		\hline \\[-1.8ex] 
		\multicolumn{4}{@{}p{36em}@{}}{\textit{Note: } The log difference in coal price between period $t$ and $s$, $ \ln (C_{t,i} / C_{s,i})$, is used as an instrument in these regressions. The sample covers all 50 US states from 2011 to 2018; outliers are removed by trimming 1\% of each variable except $\Delta_{t,s}$. The unit of observation is a set $(t,s,i)$ where $t \neq s$ are months and $i$ is a state; as discussed in the Data section, we take a random sample of 9000 observations from the data. The estimation procedure is described in the appendix.  Robust standard errors are reported in parentheses. *p$\textless$0.05, **p$\textless$0.01, ***p$\textless$0.001}  \\ 
	\end{tabular} 
\end{table}

\pagebreak
\clearpage

@@

\subsubsection{The Elasticity of Substitution between Renewable and Fossil Energy}

@@

Firstly, we let technology 1 be coal power and technology 2 be solar power; again, period $t$ represents peak hours and period $s$ represents off-peak hours. We assume that, holding prices equal, consumers prefer that 60\% of their energy arrive in period $t$ and the remaining 40\% arrive in period $s$; that is, we have $\alpha_t = 0.6$ and $\alpha_s = 0.4$. Additionally, we normalize the units of the quantities to a MWh basis. Hence, $\xi$ represents the percent of capacity utilized in each period; we assume coal uses 100\% of its capacity in both periods, while solar can access 100\% during peak hours and only 10\% during the off-peak. Finally, we set cost parameters (given in  \$ per MWh)  equal to LCOE estimates for 2023 from the \citet{EIALCOE}; specifically, we use estimates for "Solar PV" and "Coal with 30\% CCS" from Table 1b. In total, we have the following parameters.

@@

\begin{table}[t]
	\caption*{\textbf{Figure 1:} The Elasticity of Substition Between Solar and Coal} 
	\label{fig:eosnum}
	\footnotesize
	\vspace{-1em}
	\begin{tabular}{@{\extracolsep{0em}}c}
		\includegraphics[width=1\linewidth]{../code/matlab/simulation_april/fig_elasticity} \\
		\multicolumn{1}{@{\hspace{0.2in}}p{5.9in}@{}}{ \textit{Note: } The y-axis of the first plot is equivalent to $\log(X_1/X_2)$ and the x-axis of both plots is equivalent to  $\log(c_2/c_1)$. The legend in the upper subplot also applies to the lower subplot. These results were obtained using the following parameters: $\alpha_t = 0.6$, $\alpha_s = 0.4$, $\xi_1 = (1, \, 1)$, $\xi_2 = (1, \, 0.1)$, $c_1 = 104.3$, $c_2 = 60$. We plot results for various values of the intertemporal elasticity of substition for electricity consumption $\sigma$, which is used in the consumer's CES utility function. In order to numerically estimate these relationships, we first found the optimal quantities of $X$ over a range of prices $c_1^* \in (0.5\, c_1, 2 \,c_1)$. Then, we obtained estimates of the elasticity of substition by numerically differentiating $\log(X_1/X_2)$ with respect to $-log(c_1, c_2)$. That is, the elasticity of substitution between technology 1 and 2 is given by the slope of the upper subplot, and it is graphed in the lower subplot.   }     \\  
	\end{tabular}
\end{table}

We plot our numerical estimates of $e_{solar, coal}$, the elasticity of substitution between solar and coal power, in \hyperref[fig:eosnum]{Figure 1}. These results were generated by computing the optimal quantities of each technology given a range of prices. Specifically, we held the price of solar capacity constant and varied the price of coal capacity from 50\% to 200\% of its original parametrized price. Then, we numerical differentiated $\log(X_1/X_2)$ with respect to $\log(c_2/c_1)$ to estimate the elasticity of substitution between solar and coal.\footnote{Instead, we could have generated variation in $\log(c_2/c_1)$ data by holding solar at a constant price and varying the price of coal capacity. However, this does not change the results.} This numerically computed elasticity of substitution between solar and coal power, $e_{solar,coal}$, is shown in the lower subplot. We repeat this process with different values of $\sigma$, the elasticity of substitution in the consumer's utility function. Finally, we drop all observations of our numerical simulation that correspond to edge cases.

Our results show that $e$ is clearly non-linear, hence assuming a CES production structure between clean and dirty energy is not an accurate assumption. Furthermore, the relationship between log quantities and log prices becomes increasingly non-linear as $\sigma$ rises. This is somewhat intuitive; when energy consumption can be substituted across periods, the market is more sensitive to the cost efficiency of technologies and less sensitive to when they produce. For instance, suppose we had two technologies that each produced in a different period; if consumers are less willing to substitute consumption across periods, both technologies are likely to be  employed, On the other hand, when consumers are open to substitution,they will simply consume the majority of their electricity in whichever period it is most cost effective to do so. Consequently, price plays bigger role than intermittency as $\sigma$ rises. This is further evident in the lower subplots; note that the numerical estimates of  $e$ rise and become more u-shaped as $\sigma$ rises.\footnote{Recall that $\sigma < 1$ implies that  electricity consumption in different periods are complements, $\sigma = 1$ implies to a Cobb-Douglas relationship, and $\sigma > 1$ implies that electricity consumption in different periods are substitutes.}  When consumers are less concerned about when they receive their energy, they are far more open to substituting between coal and solar power. On the other hand, for lower values of $\sigma$, we find that the substitutability of coal and solar power diminishes as solar becomes progressively cheaper. To clarify, note that the x-axis of the lower subplot is $\log(c_2/c_1)$; since $c_2$ is the price of solar, cheaper solar energy implies lower values of $\log(c_2/c_1)$ which corresponds to lower values of $e$. This can be explained by $\sigma < 1$ implying a complementary relationship between electricity consumption over time. Consumers with such preferences will be relatively less sensitive to prices and more interested in spreading their electricity consumption across each period. Thus, solar is less substitutable for coal in this case, because its intermittency conflicts with consumers wanting to smooth their electricity consumption over time.

Additionally, note that we assumed earlier that the estimates of $\sigma$ on a monthly basis would be fairly close to those on a daily or hourly basis. But, our results showed estimates of $\hat{\sigma} \approx 0.5$ which differs from the literatures' estimates which fall between $0.01$ and $0.25$. Does this impact our numerical results? Alternatively, if our estimates were closer to past estimates, would there be significantly different economic implications? Based on our numerical model, the answer seems to be no. That is, the relationship between $e$ and $\log(c_2/c_1)$ in \hyperref[fig:eosnum]{Figure 1} does not significantly differ when varying $\sigma$ in $\{0.05, 0.25, 0.5\}$. Hence, even if our assumption is incorrect or past estimates are biased, our model's results do not significantly differ within this range of estimates for $\sigma$.

\subsubsection{Technical Change and the Adoption of Renewable Energy}

 \begin{table}[!t]
 	\caption*{\textbf{Figure 2:} The Market Share of Solar Energy versus Own Price} 
 	\label{fig:adoptc2}
 	\footnotesize
 	\vspace{-1em}
 	\begin{tabular}{@{\extracolsep{0em}}c}
 		\includegraphics[width=1\linewidth]{../code/matlab/simulation_april/fig_adoption_c2} \\
 		\multicolumn{1}{@{\hspace{0.2in}}p{5.9in}@{}}{ \textit{Note: } These results were obtained using the following parameters: $\alpha_t = 0.6$, $\alpha_s = 0.4$, $\xi_1 = (1, \, 1)$, $\xi_2 = (1, \, 0.1)$, $c_1 = 104.3$, $c_2 = 60$. To generate these values, we first vary the price of solar relative to its original parametrized cost $c_2$.  Then, we find the optimal capacity of solar as a fraction of the total optimal capacity of all technologies (solar and coal) for each set of prices. And, finally, we repeat this process and plot the results for different values of $\sigma$, the consumer's intertemporal elasticity of electricity consumption.   }     \\  
 	\end{tabular}
 \end{table}
